\pkg{Rgbp}: Bayesian Hierarchical Modeling and Frequentist Method Check

### 2013-04-14 jss/article.tex
just before writing baseball

                  obs.mean    n prior.mean shrinkage intv.low post.mean intv.upp post.sd
Unit w/ min(n)       0.045   67       0.02     0.834    0.011     0.024    0.042   0.008
Unit w/ median(n)    0.045  484       0.02     0.410    0.023     0.035    0.050   0.007
Unit w/ max(n)       0.020 1340       0.02     0.201    0.014     0.020    0.027   0.003
Overall Mean         0.029  517       0.02     0.438    0.015     0.025    0.037   0.006

@@

We also need a graphical summary that can give us valuable insight buried in pile of numbers and a function `\code{plot}' is exactly for this purpose.
\begin{CodeChunk}
\begin{CodeInput}
R> plot(p)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{hospital1.png}
\end{center}
\end{figure}
The regression towards the mean (RTTM) is obvious in the left-side graph; the observed sample means, empty dots on the upper line, are shrinking towards the known second-level mean (a blue vertical line at 0.02) to the different extents. Note that some hospitals' ranks have changed by shrinking much harder towards 0.02 than others. For example, the empty square symbol at the crossing point of the two left-most lines (8th and 23rd hospitals on the list above) indicates that seemingly safest hospital among 31 hospitals in terms of the observed sample mean was not actually safer than the second safest hospital. Without this hierarchical modeling, we might have made a wrong decision in choosing a hospital.
\\

@@

Then, how reliable are these intervals? Does our procedure to generate interval estimates have good repeated sampling property? The method check below will answer this question.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv <- coverage(p, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 5.5cm]{hospital2.png}
\end{center}
\end{figure}
This \code{coverage} function generated 10,000 pseudo-datasets assuming the estimated $r~(=336.37)$ as a given value. For reference, we can try any other value of $r$, for example $r=200$, by replacing above code with `\code{R> pcv <- coverage(p, A.or.r = 200, mean.PriorDist = 0.02, nsim = 10000)}.
\\

It estimated coverage probabilities by unbiased estimates (a blue dashed line) and Rao-Blackwellized estimates (a red line). Both lines are indistingishable from the horizontal line at 0.95 and the estimated overall average coverage rate is 0.951 based on both estimation methods. Finally, this method check showed that the interval estimates from the suggested multi-level modeling on this particular dataset has a very good repeated sampling property, achieving pre-specified 95\% confidence level.
\\

Morris and Christiansen (1995) also look into this ranking problem, taking shrinkage into account.
\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{school1.png}
\caption{Shrinkage and interval plots of 8 schools}
\end{center}
\end{figure}

\subsection[Unknown Second-level Mean and One Covariate]{18 Baseball Players: Unknown Prior Mean and One Covariate}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{baseball1.png}
\caption{Shrinkage and interval plots of 18 baseball players}
\end{center}
\end{figure}

>>>>

                  obs.mean    n prior.mean shrinkage low.intv post.mean
Unit w/ min(n)       0.045   67       0.02     0.834    0.011     0.024
Unit w/ median(n)    0.045  484       0.02     0.410    0.023     0.035
Unit w/ max(n)       0.020 1340       0.02     0.201    0.014     0.020
Overall Mean         0.029  517       0.02     0.438    0.015     0.025

                  upp.intv post.sd
                     0.042   0.008
                     0.050   0.007
                     0.027   0.003
                     0.037   0.006

@@

We also need a graphical summary that can give us valuable insight buried in pile of numbers and a function `\code{plot}' is exactly for this purpose.
\begin{CodeChunk}
\begin{CodeInput}
R> plot(p)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.25]{hospital1.png}
\end{center}
\end{figure}

The regression towards the mean (RTTM) is obvious in the left-side graph; the observed sample means, empty dots on the upper line, are shrinking towards the known second-level mean (a blue vertical line at 0.02) to the different extents. Note that some hospitals' ranks have changed by shrinking much harder towards 0.02 than others. For example, the empty square symbol at the crossing point of the two left-most lines (8th and 23rd hospitals on the list above) indicates that seemingly safest hospital among 31 hospitals in terms of the observed sample mean was not actually safer than the second safest hospital. Without this hierarchical modeling, we might have made a wrong decision in choosing a hospital.
\\

@@

Then, how reliable are these intervals? Does our procedure to generate interval estimates have good repeated sampling property? The following method check will answer this question. The \code{coverage} function below generates 10,000 pseudo-datasets assuming the estimated $r~(=336.37)$ as a given value. For reference, we can try any other value of $r$, for example $r=200$, by replacing below code with `\code{R> pcv <- coverage(p, A.or.r = 200, mean.PriorDist = 0.02, nsim = 10000)}.

\begin{CodeChunk}
\begin{CodeInput}
R> pcv <- coverage(p, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 5.5cm]{hospital2.png}
\end{center}
\end{figure}

It estimated coverage probabilities by unbiased estimates (a blue dashed line) and Rao-Blackwellized estimates (a red line). Both lines are indistingishable from the horizontal line at 0.95 and the estimated overall average coverage rate is 0.951 from the Rao-Blackwellized estimate and 0.95 from the unbiased estimate. Finally, this method check proved that the interval estimates from the suggested multi-level modeling on this particular dataset has a very good repeated sampling property, achieving pre-specified 95\% confidence level.
\\

For reference, \code{gbp} can give us interval estimates with different confidence level, for example 90\%,  and the appropriate code adjustment is \code{R> gbp(z, n, Alpha = 0.9, model = "pr")}.  Then, the function \code{coverage} will evaluate whether interval estimates achieve 90\% confidence level.
\\

Morris and Christiansen (1995) also looked into a similar ranking problem in hierarchical modeling, taking shrinkage into account.
\\

\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}
\begin{CodeChunk}
\begin{CodeInput}
R> g <- gbp(y, se, model = "gr")
R> g
\end{CodeInput}
\begin{CodeOutput}
Summary for whole units:

         obs.mean   se prior.mean shrinkage low.intv post.mean upp.intv post.sd
5           -1.00  9.0      8.168     0.408  -13.297     2.737   16.692   7.634
2            8.00 10.0      8.168     0.459   -7.255     8.077   23.361   7.810
7           18.00 10.0      8.168     0.459   -1.289    13.484   30.821   8.176
4            7.00 11.0      8.168     0.507   -8.780     7.592   23.602   8.257
6            1.00 11.0      8.168     0.507  -13.027     4.633   20.131   8.441
1           28.00 15.0      8.168     0.657   -2.315    14.979   38.763  10.560
3           -3.00 16.0      8.168     0.685  -17.130     4.650   22.477  10.096
8           12.00 18.0      8.168     0.734  -10.208     9.189   29.939  10.227
colMeans     8.75 12.5      8.168     0.552   -9.163     8.168   25.723   8.900
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> summary(g)
\end{CodeInput}
\begin{CodeOutput}
Main summary:

                    obs.mean   se prior.mean shrinkage low.intv post.mean
Unit w/ min(se)        -1.00  9.0      8.168     0.408  -13.297     2.737
Unit w/ median(se)1     1.00 11.0      8.168     0.507  -13.027     4.633
Unit w/ median(se)2     7.00 11.0      8.168     0.507   -8.780     7.592
Unit w/ max(se)        12.00 18.0      8.168     0.734  -10.208     9.189
Overall Mean            8.75 12.5      8.168     0.552   -9.163     8.168

                     upp.intv post.sd
                       16.692   7.634
                       20.131   8.441
                       23.602   8.257
                       29.939  10.227
                       25.723   8.900

  post.mode.alpha post.sd.alpha
1           4.768         1.139

Regression Summary:

      estimate   se z.val p.val
beta0    8.168 5.73 1.425 0.154
\end{CodeOutput}
\end{CodeChunk}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{school1.png}
\caption{Shrinkage and interval plots of 8 schools}
\end{center}
\end{figure}

\begin{CodeChunk}
\begin{CodeInput}
R> gcv <- coverage(g, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 5.5cm]{school2.png}
\end{center}
\end{figure}

\subsection[Unknown Second-level Mean and One Covariate]{18 Baseball Players: Unknown Prior Mean and One Covariate}
\begin{CodeChunk}
\begin{CodeInput}
R> b <- gbp(z, n, x, model = "br")
R> b
\end{CodeInput}
\begin{CodeOutput}
Summary for whole units:

         obs.mean  n   X1 prior.mean shrinkage low.intv post.mean upp.intv post.sd
1           0.400 45 1.00      0.310     0.715    0.248     0.335    0.429   0.046
2           0.378 45 1.00      0.310     0.715    0.244     0.329    0.420   0.045
3           0.356 45 1.00      0.310     0.715    0.240     0.323    0.411   0.044
4           0.333 45 1.00      0.310     0.715    0.236     0.316    0.403   0.043
5           0.311 45 1.00      0.310     0.715    0.230     0.310    0.396   0.042
6           0.311 45 0.00      0.233     0.715    0.179     0.256    0.341   0.041
7           0.289 45 0.00      0.233     0.715    0.175     0.249    0.331   0.040
8           0.267 45 0.00      0.233     0.715    0.171     0.243    0.323   0.039
9           0.244 45 0.00      0.233     0.715    0.166     0.237    0.315   0.038
10          0.244 45 1.00      0.310     0.715    0.210     0.291    0.379   0.043
11          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
12          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
13          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
14          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
15          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
16          0.200 45 0.00      0.233     0.715    0.155     0.224    0.302   0.038
17          0.178 45 0.00      0.233     0.715    0.148     0.218    0.297   0.038
18          0.156 45 0.00      0.233     0.715    0.140     0.211    0.292   0.039
colMeans    0.265 45 0.44      0.267     0.715    0.191     0.267    0.351   0.041
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> summary(b)
\end{CodeInput}
\begin{CodeOutput}
Main summary:

                          obs.mean  n   X1 prior.mean shrinkage low.intv
Unit w/ min(obs.mean)        0.156 45 0.00      0.233     0.715    0.140
Unit w/ median(obs.mean)1    0.244 45 0.00      0.233     0.715    0.166
Unit w/ median(obs.mean)2    0.244 45 1.00      0.310     0.715    0.210
Unit w/ max(obs.mean)        0.400 45 1.00      0.310     0.715    0.248
Overall Mean                 0.265 45 0.44      0.267     0.715    0.191

                         post.mean upp.intv post.sd
                             0.211    0.292   0.039
                             0.237    0.315   0.038
                             0.291    0.379   0.043
                             0.335    0.429   0.046
                             0.267    0.351   0.041

  post.mode.alpha post.sd.alpha
1          -4.727         0.957

Regression Summary:

      estimate    se  z.val p.val
beta0   -1.194 0.131 -9.129 0.000
beta1    0.389 0.187  2.074 0.038
\end{CodeOutput}
\end{CodeChunk}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{baseball1.png}
\caption{Shrinkage and interval plots of 18 baseball players}
\end{center}
\end{figure}

\begin{CodeChunk}
\begin{CodeInput}
R> bcv <- coverage(b, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 5.5cm]{baseball2.png}
\end{center}
\end{figure}
